---
title: "Wk2.Assignment7"
author: "Max Moro"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
data(stackloss)
```

## Week 2 Lecture Assignment

**Objective:**

- Use the stackloss data to create the following:
    - Use summary() function on data set and interpret the results.
    - Use the “str” function on the data set and interpret the results.
    - Create a scatter plot of stack.loss vs. Air.Flow and describe the relationship.

## Summary

```{r task1}
summary(stackloss)
```

The function summary shows that We have 4 columns, Air.Flow, Water.Temp, Acid.conc, and stack.loss. The functions shows then important statistics for each column, including Min, the three quartiles, and the max value. 

## Str

```{r task2}
str(stackloss)
```

the Str function shows  data structure of the stackloss data-frame. We have four columns, and they are all numeric. The data-frame contains 21 observations (rows)

## Scatter Plot

```{r task3}
plot(x=stackloss$stack.loss, y=stackloss$Air.Flow)
```

The Scatter Plot of the stackloss data shows a mostly linear and positive correlation between Stack Loss  and Air Flow variables. 
